The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-02-09 article/article.tex
revise section titles

\maketitle
\begin{abstract}

@@

@@

We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:AlwaysGoLeft}, respectively. However, we use a simplifying assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique. Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially. Section \ref{sec:analysis:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.

@@

\subsection{Analysis of the Greedy-scheme}
\label{sec:analysis:greedy}
A \emph{symmetric witness tree} of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves. Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ . The same ball may be represented by several nodes. The nodes and edges of the witness tree represent events that may occur or not. For simplicity, we assume that all events represented by a witness tree are stochastically independent.

@@

\subsubsection{Symmetric witness tree: construction}
\label{sec:constructionSymWT}
Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs. We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurance of the event. Thus, we have to show that the existance of a bin $x$ containing at least $L+4$ balls implies the existance of an activated witness tree of order $L$. We construct the witness tree as follows:
\begin{compactenum}
\item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the last ball that was inserted into the bin $x$.
\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations contain at least the same number of balls at bin $x$, we can conclude that each of the $d$ locations contain at least $L+3$ balls at the time ball $b_1$ was inserted. Thus, the $d$ children of the root node are assigned the topmost balls of these $d$ bins.
\item Consider these child nodes and continue the assignment of the previous step recursively until the leaf nodes are reached.  
\end{compactenum}

@@

\subsubsection{Symmetric witness tree: probability of activation}
\label{sec:probabilitySymWT}
In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$. The probability $p_a$ that some witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.

@@

\subsection{Analysis of the Always-go-left scheme}
\label{sec:analysis:alg}
Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the \emph{Greedy} scheme, we will use asymmetric witness trees to get an analogous upper bound for the \emph{Always-go-left} scheme.

@@

\subsection{Nondistinct balls}
\label{sec:analysis:nondistinctBalls}

>>>>

\maketitle

\begin{abstract}

@@

%\tableofcontents

@@

We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively. However, we use a simplifying assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique. Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially. Section \ref{sec:analysis:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.

@@

\subsection{Greedy}
\label{sec:analysis:greedy}
A \emph{symmetric witness tree} of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves. Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ . The same ball may be represented by several nodes. The nodes and edges of the witness tree represent events that may occur or not. For simplicity, we assume that all events represented by a witness tree are stochastically independent.

@@

Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs. We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurrence of the event. Thus, we have to show that the existence of a bin $x$ containing at least $L+4$ balls implies the existence of an activated witness tree of order $L$.

\subsubsection{Witness tree construction}
\label{sec:constructionSymWT}
We construct a witness tree for the event that some bin contains $L+4$ balls as follows:
\begin{compactenum}
\item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the last ball that was inserted into the bin $x$.
\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations contain at least the same number of balls at bin $x$, we can conclude that each of the $d$ locations contain at least $L+3$ balls at the time ball $b_1$ was inserted. Thus, the $d$ children of the root node are assigned the topmost balls of these $d$ bins.
\item Consider these child nodes and continue the assignment of the previous step recursively until the leaf nodes are reached.  
\end{compactenum}

@@

\subsubsection{Probability of activation}
\label{sec:probabilitySymWT}
In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$. The probability $p_a$ that some witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.

@@

\subsection{Always-go-left}
\label{sec:analysis:alg}
Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the \emph{Greedy} scheme, we will use asymmetric witness trees to get an analogous upper bound for the \emph{Always-go-left} scheme.

@@

\subsubsection{Witness tree construction}
\label{sec:analysis:constructionAsymWT}

\subsubsection{Probability of activation}
\label{sec:analysis:probabilityAsymWT}

\subsection{Non-distinct balls}
\label{sec:analysis:nondistinctBalls}
